---
title: "MSKIDS: Combat Harmonization of Healthy Control Volumes"
author: "Virgilio Gonzenbach & Taki Shinohara"
date: "12/18/2020"
output:
  ioslides_presentation: default
  slidy_presentation: default
bibliography: '/Users/vgonzenb/Documents/Bibtex/PennSIVE-temp.bib'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE)

library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(neuroCombat)
library(FactoMineR)
library(factoextra)
library(ggpubr)
library(sjPlot)
library(sjlabelled)

df = readRDS("data/deriv/HC_data.rds")
dict = read_xlsx("MUSE_ROI_Dict.xlsx", 1)

# Improve readibility for raw ICV
df$raw.ICV = df$X702

# Batch
site = paste(df$site, df$scanner, sep = "-") %>% as.factor()

# Drop  batch with n=7
df = df[site != "HSC-SIEMENSTIMTRIO", ]
site = site[site != "HSC-SIEMENSTIMTRIO"]
site = droplevels(site)

# Pick volumes
indexes = paste0("X", 1:300)
indexes = indexes[indexes %in% colnames(df)] # keep indexes that match a column name

# Sex Filters
M_filter = df$sex == "MALE"
F_filter = df$sex == "FEMALE"

# Tissue filters
dict_s = dict[dict$ROI_INDEX %in% 2:300, ]
VN_filter = dict_s$TISSUE_SEG == "VN"
GM_filter = dict_s$TISSUE_SEG == "GM"
WM_filter = dict_s$TISSUE_SEG == "WM"
```

## Scanners and Sites

```{r, echo=FALSE}
table(df$scanner)
table(df$site)
```

## Scanners by Sites

```{r}
table(df$scanner, df$site)
```

Excluding HSC-SIEMENSTIMTRIO (n=7) for now.

```{r, echo=FALSE}
print("Site-Scanner Pairs:");levels(site)
```

## Plot: Age

```{r}
df %>% ggplot(aes(x=site, y=age, color = site)) + geom_boxplot() 
```


## ANOVA: Age

```{r}
anova(lm(age~site, data = df))
```

Average age differs across sites.

## Plot: Sex
```{r}
# Age by sex distr
p = df %>% ggplot(aes(x= sex, color=sex, y=age)) + 
  facet_wrap(~ site) + 
  geom_boxplot()
change_palette(p, palette = "npg")
```

## Model: Sex

```{r}
table(df$site, df$sex)
table(df$site, df$sex) / colSums(table(df$sex, df$site))
chisq.test(df$site, df$sex)
```

## Layout

1. Harmonization of **intracranial volume (ICV).**  

2. Harmonization of **145** volumes  

3. ICV Normalization of 145 volumes  

Harmonization method: ComBat (@Johnson2007c, @Fortin2017c, -@Fortin2018c).

## Layout: Options 

1. Harmonization of **intracranial volume (ICV).**  
    A. As single dataset.  
    B. **By Splitting data by Males and Females (M/F).**  

2. Harmonization of **145** volumes  
    A. As single dataset.  
    B. **By splitting data by Males and Females (M/F).**  
    C. By splitting data by tissues (Tissue).  
    D. B + C.  
    
3. ICV Normalization of 145 volumes  
    A. No adjustment.  
    B. ICV Residualization: Splitting data by Males and Females (M/F).  
    C. ICV Residualization: Joint (i.e. no M/F split).  
    D. ICV as Covariate.  

## ICV (Raw)

```{r}
df %>% ggplot(aes(x = site, y = raw.ICV, color = site)) + geom_boxplot() + ylab("ICV") + facet_wrap(~sex)
```

## ICV ~ Age [PNC]
```{r}
pnc = df %>% filter(site == "PNC")
p = pnc %>% ggplot(aes(x = age, y = raw.ICV, col=sex)) + geom_point() + geom_smooth(method = "lm") + ylab("ICV") + facet_wrap(~sex)
suppressMessages(print(change_palette(p, palette = "npg"))) 
```

## Model: ICV ~ Age [PNC]

```{r}
## Std ICV -> Label ICV -> Prep Age2 -> Model
# PNC
pnc$raw.ICV_std = (pnc$raw.ICV - mean(pnc$raw.ICV)) / sd(pnc$raw.ICV)
pnc$raw.ICV_std = set_label(pnc$raw.ICV_std, "ICV")
pnc$age2 = (pnc$age - mean(pnc$age))^2
lm0 = lm(raw.ICV_std ~scale(age) + sex + scale(age2), data=pnc)

# Males
pnc_M = pnc %>% filter(sex == "MALE")
pnc_M$raw.ICV_std = (pnc_M$raw.ICV - mean(pnc_M$raw.ICV)) / sd(pnc_M$raw.ICV)
pnc_M$raw.ICV_std = set_label(pnc_M$raw.ICV_std, "ICV [M]")
pnc_M$age2 = (pnc_M$age - mean(pnc_M$age))^2
lm1 = lm(raw.ICV_std ~scale(age) + scale(age2), data=pnc_M)

# Females
pnc_F = pnc %>% filter(sex == "FEMALE")
pnc_F$raw.ICV_std = (pnc_F$raw.ICV - mean(pnc_F$raw.ICV)) / sd(pnc_F$raw.ICV)
pnc_F$raw.ICV_std = set_label(pnc_F$raw.ICV_std, "ICV [F]")
pnc_F$age2 = (pnc_F$age - mean(pnc_F$age))^2
lm2 = lm(raw.ICV_std ~scale(age) + scale(age2), data=pnc_F)

## Males: Age 2
suppressMessages(tab_model(lm0, lm1, lm2, show.ci = FALSE, p.style = "asterisk"))
```

## Model: ICV ~ Age [PNC, Males]

```{r}
pnc_M = pnc %>% filter(sex == "MALE")
pnc_M$raw.ICV_std = (pnc_M$raw.ICV - mean(pnc_M$raw.ICV)) / sd(pnc_M$raw.ICV)
pnc_M$raw.ICV_std = set_label(pnc_M$raw.ICV_std, "ICV [M]")
lm1 = lm(raw.ICV_std ~scale(age), data=pnc_M)
suppressMessages(tab_model(lm1))

pnc_F = pnc %>% filter(sex == "FEMALE")
pnc_F$raw.ICV_std = (pnc_F$raw.ICV - mean(pnc_F$raw.ICV)) / sd(pnc_F$raw.ICV)
pnc_F$raw.ICV_std = set_label(pnc_F$raw.ICV_std, "ICV [F]")
lm2 = lm(raw.ICV_std ~scale(age), data=pnc_F)
suppressMessages(tab_model(lm2))
```

## Model: ICV ~ Age [PNC, Females]

```{r}
pnc$raw.ICV_std = (pnc$raw.ICV - mean(pnc$raw.ICV)) / sd(pnc$raw.ICV)
pnc$raw.ICV_std = set_label(pnc$raw.ICV_std, "ICV")
lm0 = lm(raw.ICV_std ~scale(age), data=pnc)
suppressMessages(tab_model(lm0))
```


## site contribution
```{r}
# anova comparison
anova(lm(raw.ICV ~ sex + age + age*sex, data=df), lm(raw.ICV ~ sex + age + age*sex + site, data=df))
```

```{r include=FALSE}
# Harmonize ICV
mod = model.matrix(~ sex + age + age*sex, data = df)
resCombat = neuroCombat::neuroCombat(dat = t(df$raw.ICV),
                                     batch = site, 
                                     mod = mod,
                                     eb = FALSE,
                                     parametric=TRUE)
ICV = t(resCombat$dat.combat)
```

```{r, include=FALSE}
## Harmonize ICV per Sex
library(neuroCombat)

# Males
mod_M = model.matrix(~ age, data = df[M_filter,])
ICV_M = neuroCombat::neuroCombat(dat = t(df$raw.ICV[M_filter]), 
                                       batch = site[M_filter], 
                                       mod = mod_M,
                                       eb = FALSE,
                                       parametric=TRUE)

# Females
mod_F = model.matrix(~ age, data = df[F_filter,])
ICV_F = neuroCombat::neuroCombat(dat = t(df$raw.ICV[F_filter]), 
                                       batch = site[F_filter], 
                                       mod = mod_F,
                                       eb = FALSE,
                                       parametric=TRUE)


ICV_MF = ICV

ICV_MF[M_filter] = t(ICV_M$dat.combat)
ICV_MF[F_filter] = t(ICV_F$dat.combat)
```

## Harmonized ICV 
```{r}
# Visualize ICV All and MF
p_all = df %>% ggplot(aes(x = site, y = ICV, color = site)) + geom_boxplot() + ylab("ICV") + facet_wrap(~sex) + ggtitle("All")
p_MF = df %>% ggplot(aes(x = site, y = ICV_MF, color = site)) + geom_boxplot() + ylab("ICV") + facet_wrap(~sex) + ggtitle("M/F")

ggarrange(p_all, p_MF, nrow = 1, common.legend = TRUE, legend = "bottom")

# M/F split is the best harmonization for ICV
ICV = ICV_MF
```

## Harmonized ICV (adj. for sex)

```{r}
# Visualize ICV All and MF
p_all = df %>% ggplot(aes(x = site, y = resid(lm(ICV~ age + sex)), color = site)) + geom_boxplot() + ylab("ICV") + facet_wrap(~sex) + ggtitle("All")
p_MF = df %>% ggplot(aes(x = site, y = resid(lm(ICV_MF ~ age + sex)), color = site)) + geom_boxplot() + ylab("ICV") + facet_wrap(~sex) + ggtitle("M/F")

ggarrange(p_all, p_MF, nrow = 1, common.legend = TRUE, legend = "bottom")
```

## Harmonized ICV (adj. for age)

```{r}
# Visualize ICV All and MF
p_all = df %>% ggplot(aes(x = site, y = resid(lm(ICV~ age)), color = site)) + geom_boxplot() + ylab("ICV") + facet_wrap(~sex) + ggtitle("All")
p_MF = df %>% ggplot(aes(x = site, y = resid(lm(ICV_MF ~ age)), color = site)) + geom_boxplot() + ylab("ICV") + facet_wrap(~sex) + ggtitle("M/F")

ggarrange(p_all, p_MF, nrow = 1, common.legend = TRUE, legend = "bottom")
```


## Harmonized ICV (adj. for age, sex)

```{r}
# Visualize ICV All and MF
p_all = df %>% ggplot(aes(x = site, y = resid(lm(ICV~ age + sex)), color = site)) + geom_boxplot() + ylab("ICV") + facet_wrap(~sex) + ggtitle("All")
p_MF = df %>% ggplot(aes(x = site, y = resid(lm(ICV_MF ~ age + sex)), color = site)) + geom_boxplot() + ylab("ICV") + facet_wrap(~sex) + ggtitle("M/F")

ggarrange(p_all, p_MF, nrow = 1, common.legend = TRUE, legend = "bottom")
```

## Scanner contribution

```{r}
anova(lm(ICV ~ sex + age + sex*age, data=df), lm(ICV ~ sex + age + age*sex + site, data=df))
```

## ICV Harmonization choice

Past this point, all reference to ICV refers to ICV harmonized by splitting data by sex (M/F; Option 1B).

```{r combat, include=FALSE}
## ComBat Harmonization
mod = model.matrix(~ sex + age + age*sex + ICV, data = df)
resCombat = neuroCombat::neuroCombat(dat = t(df[, indexes]), 
                                     batch = site, 
                                     mod = mod,
                                     eb = TRUE,
                                     parametric= TRUE,
                                     ref.batch = "PNC-SIEMENSTIMTRIO")

harmon_df = df
harmon_df[, indexes] = t(resCombat$dat.combat)
```


```{r}
WM = data.frame(combat = rowSums(harmon_df[,indexes][WM_filter]))

harmon_df %>% ggplot(aes(x = site, y = WM$combat, color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("Combat")
```

```{r combat, include=FALSE}
## ComBat Harmonization
mod = model.matrix(~ sex + age + age*sex + raw.ICV, data = df)
resCombat = neuroCombat::neuroCombat(dat = t(df[, indexes]), 
                                     batch = site, 
                                     mod = mod,
                                     eb = TRUE,
                                     parametric= TRUE,
                                     ref.batch = "PNC-SIEMENSTIMTRIO")

harmon_df = df
harmon_df[, indexes] = t(resCombat$dat.combat)

WM = data.frame(combat = rowSums(harmon_df[,indexes][WM_filter]))

harmon_df %>% ggplot(aes(x = site, y = WM$combat, color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("Combat")
```


```{r}
WM = data.frame(combat = rowSums(harmon_df[,indexes][WM_filter]))

harmon_df %>% ggplot(aes(x = site, y = WM$combat, color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("Combat")
```



```{r combat_MF, include=FALSE}
## ComBat Harmonization
mod_M = model.matrix(~ age + ICV[M_filter], data = df[M_filter,])
resCombat_M = neuroCombat::neuroCombat(dat = t(df[M_filter, indexes]), 
                                     batch = site[M_filter], 
                                     mod = mod_M,
                                     eb = TRUE,
                                     parametric= TRUE,
                                     ref.batch = "PNC-SIEMENSTIMTRIO")

mod_F = model.matrix(~ age + ICV[F_filter], data = df[F_filter,])
resCombat_F = neuroCombat::neuroCombat(dat = t(df[F_filter, indexes]), 
                                     batch = site[F_filter], 
                                     mod = mod_F,
                                     eb = TRUE,
                                     parametric= TRUE,
                                     ref.batch = "PNC-SIEMENSTIMTRIO")

harmon_df_MF = df
harmon_df_MF[M_filter, indexes] = t(resCombat_M$dat.combat)
harmon_df_MF[F_filter, indexes] = t(resCombat_F$dat.combat)
```

```{r combat_tissue, include=FALSE}
vol = df[,indexes]

mod = model.matrix(~ sex + age + age*sex + ICV, data = df)
vn_combat = neuroCombat(dat = t(vol[, VN_filter]),
                        batch = site,
                        mod = mod,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
gm_combat = neuroCombat(dat = t(vol[, GM_filter]),
                        batch = site,
                        mod = mod,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
wm_combat = neuroCombat(dat = t(vol[, WM_filter]),
                        batch = site,
                        mod = mod,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
bs_combat = neuroCombat(dat = t(vol$X35), #brainstem
                        batch = site,
                        mod = mod,
                        eb = FALSE, 
                        parametric = TRUE,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )

vol[,VN_filter] = t(vn_combat$dat.combat)
vol[,GM_filter] = t(gm_combat$dat.combat)
vol[,WM_filter] = t(wm_combat$dat.combat)
vol$X35 = t(bs_combat$dat.combat)

harmon_df_tissue = df
harmon_df_tissue[,indexes] = vol
```

```{r combat_MF_tissue, include=FALSE}
# Do both MF and tissues
vol = df[,indexes]
mod_M = model.matrix(~ age + ICV[M_filter], data = df[M_filter,])
vn_combat = neuroCombat(dat = t(vol[M_filter, VN_filter]),
                        batch = site[M_filter],
                        mod = mod_M,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
gm_combat = neuroCombat(dat = t(vol[M_filter, GM_filter]),
                        batch = site[M_filter],
                        mod = mod_M,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
wm_combat = neuroCombat(dat = t(vol[M_filter, WM_filter]),
                        batch = site[M_filter],
                        mod = mod_M,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
bs_combat = neuroCombat(dat = t(vol$X35[M_filter]), #brainstem
                        batch = site[M_filter],
                        mod = mod_M,
                        eb = FALSE, 
                        parametric = TRUE,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
vol[M_filter,VN_filter] = t(vn_combat$dat.combat)
vol[M_filter,GM_filter] = t(gm_combat$dat.combat)
vol[M_filter,WM_filter] = t(wm_combat$dat.combat)
vol$X35[M_filter] = t(bs_combat$dat.combat)

# Females
mod_F = model.matrix(~ age + ICV[F_filter], data = df[F_filter,])
vn_combat = neuroCombat(dat = t(vol[F_filter, VN_filter]),
                        batch = site[F_filter],
                        mod = mod_F,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
gm_combat = neuroCombat(dat = t(vol[F_filter, GM_filter]),
                        batch = site[F_filter],
                        mod = mod_F,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
wm_combat = neuroCombat(dat = t(vol[F_filter, WM_filter]),
                        batch = site[F_filter],
                        mod = mod_F,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
bs_combat = neuroCombat(dat = t(vol$X35[F_filter]), #brainstem
                        batch = site[F_filter],
                        mod = mod_F,
                        eb = FALSE, 
                        parametric = TRUE,
                        ref.batch = "PNC-SIEMENSTIMTRIO"
                        )
vol[F_filter,VN_filter] = t(vn_combat$dat.combat)
vol[F_filter,GM_filter] = t(gm_combat$dat.combat)
vol[F_filter,WM_filter] = t(wm_combat$dat.combat)
vol$X35[F_filter] = t(bs_combat$dat.combat)

harmon_df_MF_tissue = df
harmon_df_MF_tissue[, indexes] = vol
```

## WM total

```{r}
# Compute tissue values
WM = data.frame(raw = rowSums(df[,indexes][WM_filter]), 
                combat = rowSums(harmon_df[,indexes][WM_filter]),
                MF= rowSums(harmon_df_MF[,indexes][WM_filter]),
                tissue = rowSums(harmon_df_tissue[,indexes][WM_filter]),
                MF_tissue = rowSums(harmon_df_MF_tissue[,indexes][WM_filter]))

# WM total
p1 = harmon_df %>% ggplot(aes(x = site, y = WM$combat, color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("Combat")
p2 = harmon_df_MF %>% ggplot(aes(x = site, y = WM$MF, color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("M/F")
p3 = harmon_df_tissue %>% ggplot(aes(x = site, y = WM$tissue, color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("Tissues")
p4 = harmon_df_MF_tissue %>% ggplot(aes(x = site, y = WM$MF_tissue, color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("M/F + Tissues")
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
```

## WM total (adj. for age)
```{r}
# WM total - Regressed Age
p1 = harmon_df %>% ggplot(aes(x = site, y = resid(lm(WM$combat~ age)), color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("Combat")
p2 = harmon_df_MF %>% ggplot(aes(x = site, y = resid(lm(WM$MF~ age)), color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("M/F")
p3 = harmon_df_tissue %>% ggplot(aes(x = site, y = resid(lm(WM$tissue ~age)), color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("Tissues")
p4 = harmon_df_MF_tissue %>% ggplot(aes(x = site, y = resid(lm(WM$MF_tissue ~age)), color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("M/F + Tissues")
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
```

## WM total (adj. for age + ICV)
```{r}
p1 = harmon_df %>% ggplot(aes(x = site, y = resid(lm(WM$combat~ age + ICV)), color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("Combat")
p2 = harmon_df_MF %>% ggplot(aes(x = site, y = resid(lm(WM$MF~ age + ICV)), color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("M/F")
p3 = harmon_df_tissue %>% ggplot(aes(x = site, y = resid(lm(WM$tissue ~ age + ICV)), color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("Tissues")
p4 = harmon_df_MF_tissue %>% ggplot(aes(x = site, y = resid(lm(WM$MF_tissue ~age + ICV)), color = site)) + geom_boxplot() + ylab("WM") + facet_wrap(~sex) + ggtitle("M/F + Tissues")
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
```

**M/F** lines up residualized WM volumes best.

## GM total

```{r}
GM = data.frame(raw = rowSums(df[,indexes][GM_filter]), 
                combat = rowSums(harmon_df[,indexes][GM_filter]),
                MF = rowSums(harmon_df_MF[,indexes][GM_filter]),
                tissue = rowSums(harmon_df_tissue[,indexes][GM_filter]),
                MF_tissue = rowSums(harmon_df_MF_tissue[,indexes][GM_filter]))

# GM total
p1 = harmon_df %>% ggplot(aes(x = site, y = GM$combat, color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("Combat")
p2 = harmon_df_MF %>% ggplot(aes(x = site, y = GM$MF, color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("M/F")
p3 = harmon_df_tissue %>% ggplot(aes(x = site, y = GM$tissue, color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("Tissues")
p4 = harmon_df_MF_tissue %>% ggplot(aes(x = site, y = GM$MF_tissue, color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("M/F + Tissues")
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2,common.legend = TRUE, legend = "bottom")
```

## GM total (adj. for age)

```{r}
# GM total - Regressed Age
p1 = harmon_df %>% ggplot(aes(x = site, y = resid(lm(GM$combat~ age)), color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("Combat")
p2 = harmon_df_MF %>% ggplot(aes(x = site, y = resid(lm(GM$MF~ age)), color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("M/F")
p3 = harmon_df_tissue %>% ggplot(aes(x = site, y = resid(lm(GM$tissue ~age)), color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("Tissues")
p4 = harmon_df_MF_tissue %>% ggplot(aes(x = site, y = resid(lm(GM$MF_tissue ~age)), color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("M/F + Tissues")
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
```

## GM total (adj for age + ICV)

```{r}
p1 = harmon_df %>% ggplot(aes(x = site, y = resid(lm(GM$combat~ age + ICV)), color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("Combat")
p2 = harmon_df_MF %>% ggplot(aes(x = site, y = resid(lm(GM$MF~ age + ICV)), color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("M/F")
p3 = harmon_df_tissue %>% ggplot(aes(x = site, y = resid(lm(GM$tissue ~ age + ICV)), color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("Tissues")
p4 = harmon_df_MF_tissue %>% ggplot(aes(x = site, y = resid(lm(GM$MF_tissue ~age + ICV)), color = site)) + geom_boxplot() + ylab("GM") + facet_wrap(~sex) + ggtitle("M/F + Tissues")
ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")
```

**M/F** also performs the best on residualized GM volumes.

## ROI Harmonization choice

Past this point, all reference to ROI refers refers to ROIs harmonized by splitting data by sex (M/F; Option 2B).

## ANOVAs: M/F split harmonization

For each of the 145 ROI volumes, two models were used:   
- Model 1: ROI ~ sex + age + age**x**sex (+ ICV)  
- Model 2: ROI ~ sex + age + age**x**sex (+ ICV) + scanner

Number of models for which scanner variables contributes new information is shown below.

```{r}
# Pick best harmonization
harmon_df = harmon_df_MF

## ANOVA of all regions
Ps = matrix(nrow=4, ncol=length(indexes))
colnames(Ps) = indexes

# ANOVAs Raw
l = df[,indexes] %>% lapply(function(x) anova(lm(x ~ sex + age + age*sex, data=df), 
                                              lm(x ~ sex + age + age*sex+ site, data=df)))
for(i in indexes) Ps[1,i] = l[[i]]$`Pr(>F)`[2]

# ANOVA Raw (ICV)
l = df[,indexes] %>% lapply(function(x) anova(lm(x ~ sex + age + age*sex + raw.ICV, data=df), 
                                              lm(x ~ sex + age + age*sex + raw.ICV + site, data=df)))
for(i in indexes) Ps[2,i] = l[[i]]$`Pr(>F)`[2]

# ANOVAs - Harmonized
l = harmon_df[,indexes] %>% lapply(function(x) anova(lm(x ~ sex + age + age*sex, data=df), 
                                                     lm(x ~ sex + age + age*sex + site, data=df)))
for(i in indexes) Ps[3,i] = l[[i]]$`Pr(>F)`[2]
# ANOVAs - (ICV)
l = harmon_df[,indexes] %>% lapply(function(x) anova(lm(x ~ sex + age + age*sex + ICV, data=df), 
                                                     lm(x ~ sex + age + age*sex + ICV + site, data=df)))
for(i in indexes) Ps[4,i] = l[[i]]$`Pr(>F)`[2]

p = rowSums(Ps < (0.05/145))
names(p) = c("Raw", "Raw (ICV adj)", "Combat", "Combat (ICV adj)")
p
```

## Correcting for ICV in volumes

ROIs are corrected for ICV with the formula  
$$ROI_{adj} = ROI - \beta (ICV - \overline{ICV})$$  
where $\beta$ represents the ICV-ROI slope and $\overline{ICV}$ is the average ICV of subjects in the data used for the correction [@Nordenskjold2015a].

In this presentation, ROIs are corrected for ICV separately in males and females. Effects of correcting across the data as whole is also shown.

```{r, include=FALSE}
# Pick best harmonization
harmon_df$WM = WM$MF
harmon_df$GM = GM$MF

df_1b = harmon_df
df_unadj = harmon_df

## Correction for ICV
# Get betas
b = apply(harmon_df[,c(indexes, "WM", "GM")], 2, function(x) coef(lm(x ~ ICV))[2])
b_M = apply(harmon_df[M_filter, c(indexes, "WM", "GM")], 2, function(x) coef(lm(x ~ ICV[M_filter]))[2])
b_F = apply(harmon_df[F_filter, c(indexes, "WM", "GM")], 2, function(x) coef(lm(x ~ ICV[F_filter]))[2]) 

# Adjust volumes
for(i in c(indexes, "WM", "GM")){
  harmon_df[M_filter, i] = harmon_df[M_filter, i] - b_M[i] * (ICV[M_filter] - mean(ICV[M_filter]))
  harmon_df[F_filter, i] = harmon_df[F_filter, i] - b_F[i] * (ICV[F_filter] - mean(ICV[F_filter]))
  df_1b[, i] = as.vector(df_1b[, i] - b[i] * (ICV - mean(ICV)))
  
}
```

```{r, include=FALSE}
# PCA
resPCA = FactoMineR::PCA(harmon_df[, indexes], 
                         scale.unit = FALSE, 
                         graph = FALSE,
                         ncp = 4)

resPCA_sup = FactoMineR::PCA(harmon_df[, c(indexes, "WM", "GM")], 
                         scale.unit = FALSE, 
                         graph = FALSE,
                         quanti.sup = 146:147,
                         ncp = 4)

resPCA_1b = FactoMineR::PCA(df_1b[, indexes], 
                         scale.unit = FALSE, 
                         graph = FALSE,
                         ncp = 4)

get_ctr = function(resPCA, dim){
  impt_var =  names(which(resPCA$var$contrib[,dim] > 100/145)) # which var contribute more than average: 100/145
  most_impt = sort(resPCA$var$contrib[impt_var, dim], decreasing =  TRUE)[1:6]
  index = as.numeric(gsub(".*?([0-9]+).*", "\\1", names(most_impt)))
  ctr = data.frame(ctr = unname(most_impt), index)
  
  ctr_df = cbind(contr = ctr[,1], dict[match(ctr$index, dict$ROI_INDEX),1:4])
  return(ctr_df)
}

data = data.frame(pc1 = resPCA$ind$coord[,1], 
                  pc2 = resPCA$ind$coord[,2], 
                  pc3 = resPCA$ind$coord[,3],
                  pc4 = resPCA$ind$coord[,4],
                  sex = df$sex,
                  age = df$age)
```

## PCA: Scree

(Covariance) PCA was performed with (M/F) ICV-corrected volumes. Then GM and WM totals were projected onto PCs 1 and 2.

```{r}
fviz_screeplot(resPCA, ncp=30)
```

## PCA: Loadings
```{r pca_12}
# Variables
plot.PCA(resPCA, choix = "var", axes=c(1,2), cex = 0.6, autoLab = "yes")
```

## PCA: Contributions

Dimension 1
```{r}
get_ctr(resPCA, dim = 1)
```

Dimension 2

```{r}
get_ctr(resPCA, dim = 2)
```

## PCA: Circle of Correlations
```{r}
plot.PCA(resPCA_sup, choix = "varcor", axes=c(1,2), cex = 0.5, autoLab = "yes")
```

Note: Blue lines denote supplementary projections.

## PCA: Factor Scores
```{r}
fviz_pca_ind(resPCA, choix = "ind", geom = "point", alpha.ind = 0.5, col.ind = harmon_df$sex)
```

Sex differences are retained after ICV-residualizing by sex (M/F Residualization).

## PCA: (Whole-Data ICV adj.)

```{r}
p1 = fviz_pca_ind(resPCA_1b, choix = "ind", geom = "point", alpha.ind = 0.4, col.ind = harmon_df$sex)
p2 = fviz_pca_var(resPCA_1b, choix = "var", axes=c(1,2), col.var = "steelblue",cex = 0.4)
ggarrange(p2, p1, nrow = 1)
```

Alternatively, GM loses contr. toward Dim 2 and sex differences are eliminated if ICV is residualized irrespective of sex (i.e. Joint Residualization).

## Plot: WM (unadj.) ~ Age
```{r}
p = df_unadj %>% ggplot(aes(x = age, y = WM, col=sex)) + facet_wrap(~sex) + geom_point() + geom_smooth() 
suppressMessages(print(change_palette(p, palette = "npg")))
```

## Plot: WM (M/F Res.) ~ Age

```{r}
p = harmon_df %>% ggplot(aes(x = age, y = WM, col=sex)) + facet_wrap(~sex) + geom_point() + geom_smooth()
suppressMessages(print(change_palette(p, palette = "npg")))
```

## Plot: WM (Joint Res.) ~ Age
```{r}
p = df_1b %>% ggplot(aes(x = age, y = WM, col=sex)) + facet_wrap(~sex) + geom_point() + geom_smooth()
suppressMessages(print(change_palette(p, palette = "npg")))
```

## Models: WM ~ Age

```{r}
## Standardize WM -> Label WM_std -> Model
# Unadj
df_unadj$WM_std = (df_unadj$WM - mean(df_unadj$WM)) / sd(df_unadj$WM)
df_unadj$WM_std = set_label(df_unadj$WM_std, label= "WM")
lm0 = lm(WM_std ~ scale(age) + sex + scale(age)*sex, data = df_unadj)

# M/F
harmon_df$WM_std = (harmon_df$WM - mean(harmon_df$WM)) / sd(harmon_df$WM)
harmon_df$WM_std = set_label(harmon_df$WM_std, label = "WM (M/F Res.)")
lm1 = lm(WM_std ~ scale(age) + sex + scale(age)*sex, data = harmon_df)

# Joint Res
df_1b$WM_std = (df_1b$WM - mean(df_1b$WM)) / sd(df_1b$WM)
df_1b$WM_std = set_label(df_1b$WM_std, label = "WM (Joint Res.)")
lm2 = lm(WM_std ~ scale(age) + sex + scale(age)*sex, data = df_1b)

# ICV as covariate
df_unadj4cov = df_unadj
df_unadj4cov$WM_std = (df_unadj$WM - mean(df_unadj$WM)) / sd(df_unadj$WM)
df_unadj4cov$WM_std = set_label(df_unadj4cov$WM_std, label = "WM")
lm3 = lm(WM_std ~ scale(age) + sex + scale(age)*sex + ICV, data = df_unadj4cov)

suppressMessages(tab_model(lm0, lm1, lm2, lm3, show.ci = FALSE, p.style = "asterisk"))
```

In parentheses: ICV normalization method.

## Model: WM ~ Age [Males]

```{r}
## Split M/F -> Prep Age2 -> Model
# Unadj
df_M = df_unadj %>% filter(sex == "MALE")
age2 = (df_M$age - mean(df_M$age))^2
lm0 = lm(WM_std ~ scale(age) + scale(age2), data = df_M)

# M/F
df_M = harmon_df %>% filter(sex == "MALE")
age2 = (df_M$age - mean(df_M$age))^2
lm1 = lm(WM_std ~ scale(age) + scale(age2), data = df_M)

# Joint Res
df_M = df_1b %>% filter(sex == "MALE")
age2 = (df_M$age - mean(df_M$age))^2
lm2 = lm(WM_std ~ scale(age) + scale(age2), data = df_M)

# ICV as Cov
df_M = df_unadj4cov %>% filter(sex == "MALE")
age2 = (df_M$age - mean(df_M$age))^2
lm3 = lm(WM_std ~ scale(age) + scale(age2) + ICV[M_filter], data = df_M)

suppressMessages(tab_model(lm0, lm1, lm2, lm3, show.ci = FALSE, p.style = "asterisk"))
```

## Model: WM ~ Age [Females]

```{r}
## Split M/F -> Prep Age2 -> Model
# Unadj
df_F = df_unadj %>% filter(sex == "FEMALE")
age2 = (df_F$age - mean(df_F$age))^2
lm0 = lm(WM_std ~ scale(age) + scale(age2), data = df_F)

# M/F
df_F = harmon_df %>% filter(sex == "FEMALE")
age2 = (df_F$age - mean(df_F$age))^2
lm1 = lm(WM_std ~ scale(age) + scale(age2), data = df_F)

# Joint Res
df_F = df_1b %>% filter(sex == "FEMALE")
age2 = (df_F$age - mean(df_F$age))^2
lm2 = lm(WM_std ~ scale(age) + scale(age2), data = df_F)

# ICV as Cov
df_F = df_unadj4cov %>% filter(sex == "FEMALE")
age2 = (df_F$age - mean(df_F$age))^2
lm3 = lm(WM_std ~ scale(age) + scale(age2) + ICV[F_filter], data = df_F)

suppressMessages(tab_model(lm0, lm1, lm2, lm3, show.ci = FALSE, p.style = "asterisk"))
```

## Plot: GM (unadj.) ~ Age 
```{r}
p = df_unadj %>% ggplot(aes(x = age, y = GM)) + facet_wrap(~sex) + geom_point() + geom_smooth() 
suppressMessages(print(change_palette(p, palette = "npg")))
```

## Plot: GM (M/F Res.) ~ Age 

```{r}
p = harmon_df %>% ggplot(aes(x = age, y = GM)) + facet_wrap(~sex) + geom_point() + geom_smooth()
suppressMessages(print(change_palette(p, palette = "npg")))
```

## Plot: GM (Joint Res.) ~ Age
```{r}
p = df_1b %>% ggplot(aes(x = age, y = GM)) + facet_wrap(~sex) + geom_point() + geom_smooth()
suppressMessages(print(change_palette(p, palette = "npg")))
```

## Model: GM ~ Age

```{r}
## Standardize WM -> Label WM_std -> Model
# Unadj
df_unadj$GM_std = (df_unadj$GM - mean(df_unadj$GM)) / sd(df_unadj$GM)
df_unadj$GM_std = set_label(df_unadj$GM_std, label= "GM")
lm0 = lm(GM_std ~ scale(age) + sex + scale(age)*sex, data = df_unadj)

# M/F
harmon_df$GM_std = (harmon_df$GM - mean(harmon_df$GM)) / sd(harmon_df$GM)
harmon_df$GM_std = set_label(harmon_df$GM_std, label = "GM (M/F Res.)")
lm1 = lm(GM_std ~ scale(age) + sex + scale(age)*sex, data = harmon_df)

# 1B
df_1b$GM_std = (df_1b$GM - mean(df_1b$GM)) / sd(df_1b$GM)
df_1b$GM_std = set_label(df_1b$GM_std, label = "GM (Joint Res.)")
lm2 = lm(GM_std ~ scale(age) + sex + scale(age)*sex, data = df_1b)

# ICV as covariate
df_unadj4cov = df_unadj
df_unadj4cov$GM_std = (df_unadj$GM - mean(df_unadj$GM)) / sd(df_unadj$GM)
df_unadj4cov$GM_std = set_label(df_unadj4cov$GM_std, label = "GM (Joint Cov.)")
lm3 = lm(GM_std ~ scale(age) + sex + scale(age)*sex + ICV, data = df_unadj4cov)

suppressMessages(tab_model(lm0, lm1, lm2, lm3, show.ci = FALSE, p.style = "asterisk"))
```

## Model: GM ~ Age [Males]

```{r}
## Split M/F -> Prep Age2 -> Model
# Unadj
df_M = df_unadj %>% filter(sex == "MALE")
age2 = (df_M$age - mean(df_M$age))^2
lm0 = lm(GM_std ~ scale(age) + scale(age2), data = df_M)

# M/F
df_M = harmon_df %>% filter(sex == "MALE")
age2 = (df_M$age - mean(df_M$age))^2
lm1 = lm(GM_std ~ scale(age) + scale(age2), data = df_M)

# Joint Res
df_M = df_1b %>% filter(sex == "MALE")
age2 = (df_M$age - mean(df_M$age))^2
lm2 = lm(GM_std ~ scale(age) + scale(age2), data = df_M)

# ICV as Cov
df_M = df_unadj4cov %>% filter(sex == "MALE")
age2 = (df_M$age - mean(df_M$age))^2
lm3 = lm(GM_std ~ scale(age) + scale(age2) + ICV[M_filter], data = df_M)

suppressMessages(tab_model(lm0, lm1, lm2, lm3, show.ci = FALSE, p.style = "asterisk"))
```

## Model: GM ~ Age [Females]

```{r}
## Split M/F -> Prep Age2 -> Model
# Unadj
df_F = df_unadj %>% filter(sex == "FEMALE")
age2 = (df_F$age - mean(df_F$age))^2
lm0 = lm(GM_std ~ scale(age) + scale(age2), data = df_F)

# M/F
df_F = harmon_df %>% filter(sex == "FEMALE")
age2 = (df_F$age - mean(df_F$age))^2
lm1 = lm(GM_std ~ scale(age) + scale(age2), data = df_F)

# Joint Res
df_F = df_1b %>% filter(sex == "FEMALE")
age2 = (df_F$age - mean(df_F$age))^2
lm2 = lm(GM_std ~ scale(age) + scale(age2), data = df_F)

# ICV as Cov
df_F = df_unadj4cov %>% filter(sex == "FEMALE")
age2 = (df_F$age - mean(df_F$age))^2
lm3 = lm(GM_std ~ scale(age) + scale(age2) + ICV[F_filter], data = df_F)

suppressMessages(tab_model(lm0, lm1, lm2, lm3, show.ci = FALSE, p.style = "asterisk"))
```

## References

<Insert references for Combat and ICV-normalization.>