---
title: "MSKIDS: Exploratory Analysis and ComBat Harmonization with Healthy Controls"
author: "Virgilio Gonzenbach @ PennSIVE"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      cache = TRUE, 
                      cache.path = "cache/", 
                      fig.path = "report/fig/")
```

```{r data, include=FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(FactoMineR)
library(neuroCombat)
library(ggpubr)

df = readRDS("data/deriv/HC_data.rds")
dict = read_xlsx("MUSE_ROI_Dict.xlsx", 1)
```

## Batch Variables

```{r, echo=FALSE}
table(df$scanner)
table(df$site)
```

## Scanners by sites

```{r}
table(df$scanner, df$site)
```

Excluding HSC-SIEMENSTIMTRIO (n=7) for now.

```{r, echo=FALSE}
# Create batch variable
batch = paste(df$site, df$scanner, sep = "-") %>% as.factor()

df = df[batch != "HSC-SIEMENSTIMTRIO", ]
batch = batch[batch != "HSC-SIEMENSTIMTRIO"]
batch = droplevels(batch)
print("Batches:");levels(batch)
```


## Confounding across sites

### Age

```{r}
df %>% ggplot(aes(x=site, y=age, color = site)) + geom_boxplot() 

anova(lm(age~site, data = df))
```

Age is confounded across sites.

\pagebreak 

### Sex

```{r}
table(df$site, df$sex)
table(df$site, df$sex) / colSums(table(df$sex, df$site))
chisq.test(df$site, df$sex)
```

Sex slightly confounded.

```{r}
# Age by sex distr
p = df %>% ggplot(aes(x= sex, color=sex, y=age)) + 
  facet_wrap(~ site) + 
  geom_boxplot()

change_palette(p, palette = "npg")
```

\pagebreak

## ROI volumes

```{r}
indexes = paste0("X", 1:300)
indexes = indexes[indexes %in% colnames(df)] # keep indexes that match a column name

vol = data.frame(batch, df[,indexes])

# ANOVAs
Ps = matrix(nrow=2, ncol=length(indexes))
colnames(Ps) = indexes

l = vol[,-1] %>% lapply(function(x) anova(lm(x ~ df$sex + df$age), lm(x~df$sex + df$age + batch)))
for(i in indexes) Ps[1,i] = l[[i]]$`Pr(>F)`[2]
```

```{r, include=FALSE}
## ComBat Harmonization
library(neuroCombat)

mod = model.matrix(~age+sex, data = df)
resCombat = neuroCombat::neuroCombat(dat = t(vol[,-1]), 
                                     batch = batch, 
                                     mod = mod,
                                     eb = TRUE,
                                     parametric=TRUE)

harmon_df = data.frame(batch = batch, t(resCombat$dat.combat))

# ANOVAs - Unadj
l = harmon_df[,-1] %>% lapply(function(x) anova(lm(x ~ df$sex + df$age), lm(x ~ batch + df$sex + df$age)))
for(i in indexes) Ps[2,i] = l[[i]]$`Pr(>F)`[2]
```

### Number of significant ANOVA after accounting for sites

```{r}
p = rowSums(Ps < (0.05/145))
names(p) = c("Raw", "Combat")
p
```


```{r, include=FALSE}
# Sort ROI by volume
mean_vol = colMeans(vol[,-1])
quantiles = mean_vol %>% quantile(seq(0, 1, 0.1))

group_index = list(vector("character"), 
                   vector("character"), 
                   vector("character"), 
                   vector("character"), 
                   vector("character"),
                   vector("character"), 
                   vector("character"), 
                   vector("character"), 
                   vector("character"), 
                   vector("character"))

for(i in 2:length(quantiles)){
  if(i != length(quantiles)){
    group_index[[i-1]] = names(which(mean_vol >= quantiles[i-1] & mean_vol < quantiles[i]))
  }else{
    group_index[[i-1]] = names(which(mean_vol >= quantiles[i-1] & mean_vol <= quantiles[i]))
  }
}
```


```{r, echo=FALSE}
## Boxplots
plot_roi = function(dat1, dat2, group_index){
  plots = list()
  j = 1
  for(i in group_index){
    p1 = dat1 %>% select(batch, i) %>% 
      gather(region, volume, -batch) %>% 
      ggplot(aes(x=region, y=volume, color=batch)) + geom_boxplot() + 
      theme(axis.text.x=element_text(size=7, angle=45),
        axis.ticks.x=element_blank()) + ggtitle("Raw")
    
    p2 = dat2 %>% select(batch, i) %>% 
      gather(region, volume, -batch) %>% 
      ggplot(aes(x=region, y=volume, color=batch)) + geom_boxplot() + 
      theme(axis.text.x=element_text(size=7, angle=45), 
        axis.ticks.x=element_blank()) + ggtitle("Combat")
    
    
    plots[[j]] = ggarrange(p1, p2, ncol = 1, common.legend = TRUE, legend = "bottom")
    j = j + 1
  }
  return(plots)
}

roi_plots = plot_roi(vol, harmon_df, group_index)
```

### ROI plots
```{r}
roi_plots
```

## (Covariance) PCA

PC1: total volume (w/ higher contribution from WM areas)
PC2: GM/WM
PC3: Cerebellum
PC4: L. Ventricles vs Frontal Lobes

```{r, include=FALSE}
# Covariance PCA
resPCA = FactoMineR::PCA(vol, 
                scale.unit = FALSE, 
                ncp = 6, 
                quali.sup = 1, 
                graph = FALSE)

# PCA Combat
resPCA_Combat = FactoMineR::PCA(harmon_df, 
                                scale.unit = FALSE, 
                                ncp = 6, 
                                quali.sup = 1,
                                graph = FALSE)

means = as.data.frame(cbind(batch, resPCA$ind$coor)) %>% group_by(batch) %>% summarize_all(mean)
means_Combat = as.data.frame(cbind(batch, resPCA_Combat$ind$coor)) %>% group_by(batch) %>% summarize_all(mean)
```

```{r}
resPCA_Combat$eig[1:4,]
```

```{r}
# Factor scores (Variables)
plot.PCA(resPCA_Combat, axes=c(1,2),choix = "var", cex = 1, autoLab = "yes")
```


```{r, echo=FALSE}
impt_var =  names(which(resPCA_Combat$var$contrib[,1] > 1/145))
most_impt = sort(resPCA_Combat$var$contrib[impt_var, 1], decreasing =  TRUE)[1:10]

index = as.numeric(gsub(".*?([0-9]+).*", "\\1", names(most_impt)))
ctr = data.frame(ctr = unname(most_impt), index)

cbind(contr = ctr[,1], dict[match(ctr$index, dict$ROI_INDEX),1:4])
```

```{r, echo=FALSE}
impt_var =  names(which(resPCA_Combat$var$contrib[,2] > 1/145))
most_impt = sort(resPCA_Combat$var$contrib[impt_var, 2], decreasing =  TRUE)[1:10]

index = as.numeric(gsub(".*?([0-9]+).*", "\\1", names(most_impt)))
ctr = data.frame(ctr = unname(most_impt), index)

cbind(contr = ctr[,1], dict[match(ctr$index, dict$ROI_INDEX),1:4])
```

```{r}
plot.PCA(resPCA_Combat, axes=c(3,4),choix = "var", cex = 1, autoLab = "yes")
```

```{r, echo=FALSE}
impt_var =  names(which(resPCA_Combat$var$contrib[,3] > 1/145))
most_impt = sort(resPCA_Combat$var$contrib[impt_var, 3], decreasing =  TRUE)[1:10]

index = as.numeric(gsub(".*?([0-9]+).*", "\\1", names(most_impt)))
ctr = data.frame(ctr = unname(most_impt), index)

cbind(contr = ctr[,1], dict[match(ctr$index, dict$ROI_INDEX),1:4])
```

```{r, echo=FALSE}
impt_var =  names(which(resPCA_Combat$var$contrib[,4] > 1/145))
most_impt = sort(resPCA_Combat$var$contrib[impt_var, 4], decreasing =  TRUE)[1:10]

index = as.numeric(gsub(".*?([0-9]+).*", "\\1", names(most_impt)))
ctr = data.frame(ctr = unname(most_impt), index)

cbind(contr = ctr[,1], dict[match(ctr$index, dict$ROI_INDEX),1:4])
```


```{r, include=FALSE}
# Dataframes: Raw and ComBat
dat = data.frame(pc1 = resPCA$ind$coord[,1], 
                 pc2 = resPCA$ind$coord[,2], 
                 pc3 = resPCA$ind$coord[,3],
                 pc4 = resPCA$ind$coord[,4],
                 sex = df$sex, 
                 age = df$age, 
                 batch = batch)

dat_combat = data.frame(pc1 = resPCA_Combat$ind$coord[,1], 
                        pc2 = resPCA_Combat$ind$coord[,2], 
                        pc3 = resPCA_Combat$ind$coord[,3], 
                        pc4 = resPCA_Combat$ind$coord[,4],
                        sex = df$sex, 
                        age = df$age, 
                        batch = batch)
```

### PC1

```{r, echo=FALSE}
# Histograms: Dims per Site
dim = 1
pc_raw = data.frame(resPCA$ind$coor, batch) %>% ggplot(aes(x=dat[,dim], color=batch)) + 
  geom_histogram(alpha=0.2, position="identity", bins = 18) + 
  geom_vline(xintercept = as.matrix(means[,dim+1]), 
             color=c("#F8766D", "#00BA38", "#619CFF"),
             linetype="dashed") +
  ggtitle("Raw Dimension 1 Scores") + xlab("Dimension 1")


pc_combat = data.frame(resPCA_Combat$ind$coor, batch) %>% ggplot(aes(x=dat_combat[,dim], color=batch)) + 
  geom_histogram(alpha=0.2, position="identity", bins = 18) + 
  geom_vline(xintercept = as.matrix(means_Combat[,dim+1]), color=c("#F8766D", "#00BA38", "#619CFF"),
             linetype="dashed") +
  ggtitle("Harmonized Dimension 1 Scores") + xlab("Dimension 1")

ggarrange(pc_raw, pc_combat, ncol = 1, common.legend = TRUE, legend = "bottom", align = "hv")

print("RAW");   summary(glm(scale(pc1) ~ batch + age + sex, data = dat))
print("COMBAT");summary(glm(scale(pc1) ~ batch + age + sex, data = dat_combat))
```

### PC2

```{r, echo=FALSE}
# Histograms: Dims per Site
dim = 2
pc_raw = data.frame(resPCA$ind$coor, batch) %>% ggplot(aes(x=resPCA$ind$coord[,dim], color=batch)) + 
  geom_histogram(alpha=0.2, position="identity", bins = 18) + 
  geom_vline(xintercept = as.matrix(means[,dim+1]), 
             color=c("#F8766D", "#00BA38", "#619CFF"),
             linetype="dashed") +
  ggtitle("Raw Dimension 2 Scores") + xlab("Dimension 2")


pc_combat = data.frame(resPCA_Combat$ind$coor, batch) %>% ggplot(aes(x=resPCA_Combat$ind$coord[,dim], color=batch)) + 
  geom_histogram(alpha=0.2, position="identity", bins = 18) + 
  geom_vline(xintercept = as.matrix(means_Combat[,dim+1]), color=c("#F8766D", "#00BA38", "#619CFF"),
             linetype="dashed") +
  ggtitle("Harmonized Dimension 2 Scores") + xlab("Dimension 2")

ggarrange(pc_raw, pc_combat, ncol = 1, common.legend = TRUE, legend = "bottom")

print("RAW");   summary(glm(scale(pc2) ~ batch + age + sex, data = dat))
print("COMBAT");summary(glm(scale(pc2) ~ batch + age + sex, data = dat_combat))
```

###PC3
```{r, echo=FALSE}
# Histograms: Dims per Site
dim = 3
pc_raw = data.frame(resPCA$ind$coor, batch) %>% ggplot(aes(x=resPCA$ind$coord[,dim], color=batch)) + 
  geom_histogram(alpha=0.2, position="identity", bins = 18) + 
  geom_vline(xintercept = as.matrix(means[,dim+1]), 
             color=c("#F8766D", "#00BA38", "#619CFF"),
             linetype="dashed") +
  ggtitle("Raw Dimension 3 Scores") + xlab("Dimension 3")


pc_combat = data.frame(resPCA_Combat$ind$coor, batch) %>% ggplot(aes(x=resPCA_Combat$ind$coord[,dim], color=batch)) + 
  geom_histogram(alpha=0.2, position="identity", bins = 18) + 
  geom_vline(xintercept = as.matrix(means_Combat[,dim+1]), color=c("#F8766D", "#00BA38", "#619CFF"),
             linetype="dashed") +
  ggtitle("Harmonized Dimension 3 Scores") + xlab("Dimension 3")

ggarrange(pc_raw, pc_combat, ncol = 1, common.legend = TRUE, legend = "bottom")

print("RAW");   summary(glm(scale(pc3) ~ batch + age + sex, data = dat))
print("COMBAT");summary(glm(scale(pc3) ~ batch + age + sex, data = dat_combat))
```

### PC4

```{r, echo=FALSE}
# Histograms: Dims per Site
dim = 4
pc_raw = data.frame(resPCA$ind$coor, batch) %>% ggplot(aes(x=resPCA$ind$coord[,dim], color=batch)) + 
  geom_histogram(alpha=0.2, position="identity", bins = 18) + 
  geom_vline(xintercept = as.matrix(means[,dim+1]), 
             color=c("#F8766D", "#00BA38", "#619CFF"),
             linetype="dashed") +
  ggtitle("Raw Dimension 4 Scores") + xlab("Dimension 4")

pc_combat = data.frame(resPCA_Combat$ind$coor, batch) %>% ggplot(aes(x=resPCA_Combat$ind$coord[,dim], color=batch)) + 
  geom_histogram(alpha=0.2, position="identity", bins = 18) + 
  geom_vline(xintercept = as.matrix(means_Combat[,dim+1]), color=c("#F8766D", "#00BA38", "#619CFF"),
             linetype="dashed") +
  ggtitle("Harmonized Dimension 4 Scores") + xlab("Dimension 4")

ggarrange(pc_raw, pc_combat, ncol = 1, common.legend = TRUE, legend = "bottom")

print("RAW");   summary(glm(scale(pc4) ~ batch + age + sex, data = dat))
print("COMBAT");summary(glm(scale(pc4) ~ batch + age + sex, data = dat_combat))
```


\pagebreak

## PCs ~ Age

### PC1

```{r, echo=FALSE}
# Raw
dat %>% 
  ggplot(aes(x=age, y=pc1,col=batch)) + geom_point() + ggtitle("Raw") + geom_smooth(method="lm") + facet_wrap(~sex)

print("RAW");   summary(glm(scale(pc1) ~ batch + scale(age) + sex, data = dat))
```

```{r, echo=FALSE}
# Combat
dat_combat %>% ggplot(aes(x=age, y=pc1,col=batch)) + geom_point() + facet_wrap(~sex) + ggtitle("Combat") + geom_smooth(method="lm")

print("COMBAT");summary(glm(scale(pc1) ~ batch + scale(age) + sex, data = dat_combat))

```

### PC2

```{r, echo=FALSE}
# PC2 and age
dat %>% ggplot(aes(x=age, y=pc2,col=batch)) + geom_point() + ggtitle("Raw") + facet_wrap(~sex) + geom_smooth(method="lm")

print("RAW");   summary(glm(scale(pc2) ~ batch + scale(age) + sex, data = dat))
```

```{r, echo=FALSE}
dat_combat %>% ggplot(aes(x=age, y=pc2,col=batch)) + geom_point() + ggtitle("Combat") + geom_smooth(method="lm") + facet_wrap(~sex)

print("COMBAT");summary(glm(scale(pc2) ~ batch + scale(age) + sex, data = dat_combat))

```

### PC3

```{r, echo=FALSE}
# PC3 and age
dat %>% ggplot(aes(x=age, y=pc3,col=batch)) + geom_point() + ggtitle("Raw") + geom_smooth(method="lm")

print("RAW");   summary(glm(scale(pc3) ~ batch + scale(age) + sex, data = dat))
```

```{r, echo=FALSE}
dat_combat %>% ggplot(aes(x=age, y=pc3,col=batch)) + geom_point() + ggtitle("Combat") + geom_smooth(method="lm")

print("COMBAT");summary(glm(scale(pc3) ~ batch + scale(age) + sex, data = dat_combat))
```

### PC4

```{r, echo=FALSE}
# PC4 and age
dat %>% ggplot(aes(x=age, y=pc4,col=batch)) + geom_point() + ggtitle("Raw") + geom_smooth(method="lm")

print("RAW");   summary(glm(scale(pc4) ~ batch + scale(age) + sex, data = dat))
```

```{r, echo=FALSE}
dat_combat %>% ggplot(aes(x=age, y=pc4,col=batch)) + geom_point() + ggtitle("Combat") + geom_smooth(method="lm")

print("COMBAT");summary(glm(scale(pc4) ~ batch + scale(age) + sex, data = dat_combat))
```

\pagebreak

## PCs ~ Sex

### PC1
```{r, echo=FALSE}
# PC1 and sex
p1 = dat %>% ggplot(aes(x=sex, y=scale(pc1), color=sex)) + geom_boxplot() + ggtitle("Raw") + facet_wrap(~batch) 
p1 = p1 %>% change_palette(palette="png")

p2 = dat_combat %>% ggplot(aes(x=sex, y=scale(pc1), color=sex)) + geom_boxplot() + ggtitle("Combat") + facet_wrap(~batch)
p2 = p2 %>% change_palette(palette="png")

ggarrange(p1, p2, ncol = 1, common.legend = TRUE, legend = FALSE)
```

```{r, echo=FALSE}
print("RAW");   summary(glm(scale(pc1) ~ batch + sex + scale(age), data = dat))
print("COMBAT");summary(glm(scale(pc1) ~ batch + sex + scale(age), data = dat_combat))
```


### PC2
```{r, echo=FALSE}
# PC2 and sex
p1 = dat %>% ggplot(aes(x=sex, y=scale(pc2), color=sex)) + geom_boxplot() + ggtitle("Raw") + facet_wrap(~batch) 
p1 = p1 %>% change_palette(palette="png")

p2 = dat_combat %>% ggplot(aes(x=sex, y=scale(pc2), color=sex)) + geom_boxplot() + ggtitle("Combat") + facet_wrap(~batch)
p2 = p2 %>% change_palette(palette="png")

ggarrange(p1, p2, ncol = 1, common.legend = TRUE, legend = FALSE)
```

```{r, echo=FALSE}
print("RAW");   summary(glm(scale(pc2) ~ batch + sex + scale(age), data = dat))
print("COMBAT");summary(glm(scale(pc2) ~ batch + sex + scale(age), data = dat_combat))
```

### PC3
```{r, echo=FALSE}
# PC3 and sex
p1 = dat %>% ggplot(aes(x=sex, y=scale(pc3), color=sex)) + geom_boxplot() + ggtitle("Raw") + facet_wrap(~batch) 
p1 = p1 %>% change_palette(palette="png")

p2 = dat_combat %>% ggplot(aes(x=sex, y=scale(pc3), color=sex)) + geom_boxplot() + ggtitle("Combat") + facet_wrap(~batch)
p2 = p2 %>% change_palette(palette="png")

ggarrange(p1, p2, ncol = 1, common.legend = TRUE, legend = FALSE)
```

```{r, echo=FALSE}
print("RAW");   summary(glm(scale(pc3) ~ batch + sex + scale(age), data = dat))
print("COMBAT");summary(glm(scale(pc3) ~ batch + sex + scale(age), data = dat_combat))
```

### PC4
```{r, echo=FALSE}
# PC4 and sex
p1 = dat %>% ggplot(aes(x=sex, y=scale(pc4), color=sex)) + geom_boxplot() + ggtitle("Raw") + facet_wrap(~batch) 
p1 = p1 %>% change_palette(palette="png")

p2 = dat_combat %>% ggplot(aes(x=sex, y=scale(pc4), color=sex)) + geom_boxplot() + ggtitle("Combat") + facet_wrap(~batch)
p2 = p2 %>% change_palette(palette="png")

ggarrange(p1, p2, ncol = 1, common.legend = TRUE, legend = FALSE)
```

```{r, echo=FALSE}
print("RAW");   summary(glm(scale(pc4) ~ batch + sex + scale(age), data = dat))
print("COMBAT");summary(glm(scale(pc4) ~ batch + sex + scale(age), data = dat_combat))
```

```

