---
title: "MSKIDS Scanner Harmonization"
author: "Virgilio Gonzenbach"
date: "3/22/2021"
output: 
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "beaver"
    fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      cache = TRUE, 
                      cache.path = 'cache/GAMs/', 
                      fig.path = "figures/GAMs/")
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r, warning=FALSE, results='hide', message=FALSE}
library(dplyr)
library(ggplot2) 
library(mgcv)
library(readxl)
library(psych)
library(kableExtra)
library(stargazer)
library(pander)
library(ggpubr)
library(reshape2)
library(readr)

df = readRDS("data/deriv/HC_data.rds")
dict = read_xlsx("data/MUSE_ROI_Dict.xlsx", 1)
dict = dict[dict$ROI_INDEX %in% 2:300, ]

# Load data
HC_raw = read.csv('data/deriv/HC_data.csv')
HC_raw = HC_raw %>% mutate(MS = 0, .after = age)
MS_raw = read.csv('data/deriv/MS_data.csv')
MS_raw = MS_raw %>% mutate(MS = 1, .after = age)

PNC = HC_raw %>% filter(site == "PNC")
HC_raw = HC_raw %>% filter(site != "PNC")

HC_GAM = read.csv('data/deriv/HC_adj_ComBat-GAM_from-HC+MS_split-MF_eb-True.csv', stringsAsFactors = TRUE)
MS_GAM = read.csv('data/deriv/MS_adj_ComBat-GAM_from-HC+MS_split-MF_eb-True.csv', stringsAsFactors = TRUE)

list_df = list(HC_raw = HC_raw, MS_raw = MS_raw, PNC = PNC, HC_GAM = HC_GAM, MS_GAM = MS_GAM)

# Rename ICV for readibility
for(i in 1:length(list_df)){
   ICV_index = which(colnames(list_df[[i]]) == "X702")
   colnames(list_df[[i]])[ICV_index] = "ICV"
} 

prep_rois = function(df){
   
   WM_indexes = paste0("X", dict$ROI_INDEX[dict$TISSUE_SEG == "WM"])
   GM_indexes = paste0("X", dict$ROI_INDEX[dict$TISSUE_SEG == "GM"])
   VN_indexes = paste0("X", dict$ROI_INDEX[dict$TISSUE_SEG == "VN"])

   #n Load 
   sig_rois = readRDS('data/non_linear_rois.rds') # ROIs that exhibit nonlinear effects with age
   
   df = df %>% mutate(WM = rowSums(across(WM_indexes)),
                      GM = rowSums(across(GM_indexes)),
                      CSF = rowSums(across(VN_indexes)),
                      Brain.Stem = rowSums(across(sig_rois[1])),
                      Cerebellum.WM = rowSums(across(sig_rois[2:3])),
                      Ventral.DC = rowSums(across(sig_rois[4:5])),
                      Fornix = rowSums(across(sig_rois[6:7])),
                      Ant.Limb.Int.Capsule = rowSums(across(sig_rois[8:9])))
   
   # Prepare covariates: square age
   df = df %>% mutate(age2 = age - mean(age)^2, .after = age)
   
   df$sex = as.factor(df$sex)
   return(df)
}

list_df = list_df %>% lapply(prep_rois)

HC_raw = list_df[[1]]
MS_raw = list_df[[2]]
PNC    = list_df[[3]]
HC_GAM = list_df[[4]]
MS_GAM = list_df[[5]]

# prepare target variables
roi_small = paste0("X", dict$ROI_INDEX)
roi_big = colnames(list_df[[1]])[153:160]
```

## Overview

- MSKIDS data summary
- PNC analysis: ROIs [Males vs Females], GAMs
- Harmonization of scanner effects with ComBat-GAM

# MSKIDS data summary

## MSKIDS: All participants

Participants: `r nrow(rbind(HC_raw, MS_raw))`

Per scanner:
```{r}
rbind(HC_raw, MS_raw) %>% group_by(site) %>% count() %>% kable()
```

Per sex and site:
```{r}
rbind(HC_raw, MS_raw) %>% group_by(sex) %>% count(site) %>% kable()
```

Totals: Females = `r sum(rbind(HC_raw, MS_raw)$sex == "FEMALE")`; Males = `r sum(rbind(HC_raw, MS_raw)$sex == "MALE")`

## MSKIDS: HC only

Participants: `r nrow(HC_raw)`

Per site:
```{r}
HC_raw %>% group_by(site) %>% count() %>% kable()
```

Per sex and site:
```{r}
HC_raw %>% group_by(sex) %>% count(site) %>% kable()
```

Totals: Females = `r sum(HC_raw$sex == "FEMALE")`; Males = `r sum(HC_raw$sex == "MALE")`

## MSKIDS: MS only

Participants: `r nrow(MS_raw)`

Per site:
```{r}
MS_raw %>% group_by(site) %>% count() %>% kable()
```

Per sex and site:
```{r}
MS_raw %>% group_by(sex) %>% count(site) %>% kable()
```

Totals: Females = `r sum(MS_raw$sex == "FEMALE")`; Males = `r sum(MS_raw$sex == "MALE")`

## MSKIDS: Age [All participants]
```{r}
rbind(HC_raw, MS_raw) %>% ggplot(aes(x=site, col=site, y=age)) + geom_boxplot() + facet_wrap(~sex) + xlab("Site") + ylab("Age") + ggtitle("Age of MSKIDS participants") + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank())
```

## MSKIDS: Age [HC only]

```{r}
HC_raw %>% ggplot(aes(x=site, col=site, y=age)) + geom_boxplot() + facet_wrap(~sex) + xlab("Site") + ylab("Age") + ggtitle("Age of MSKIDS Healthy Controls") + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank())
```

## MSKIDS: Age [MS only]

```{r}
MS_raw %>% ggplot(aes(x=site, col=site, y=age)) + geom_boxplot() + facet_wrap(~sex) + xlab("Site") + ylab("Age") + ggtitle("Age of MSKIDS MS Patients") + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank())
```

# PNC Analysis

## Overview

Objective: To inform harmonization approach by  
- Determining Age-ROI, ICV-Age, ICV-ROI relationships in Males and Females  
- Ascertaining non-linear age trends in select ROIs

## Age by Sex

```{r, fig.width=7, fig.height=4}
PNC %>% ggplot(aes(x=sex, col=sex, y=age)) + geom_boxplot() + xlab("Sex") + ylab("Age") + ggtitle("Age of the Philadelphia Neurodevelopment Cohort") + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank())
```

```{r}
sex_count = PNC %>% group_by(sex) %>% count()
```
Females: `r sex_count$n[sex_count$sex == "FEMALE"]` 
Males : `r sex_count$n[sex_count$sex == "MALE"]` 

## Age descriptives
Full dataset:   
```{r}
descr = describe(PNC$age)[,c(2:5, 8:9),]
row.names(descr) = "age"
descr %>% pander()
```

Males:
```{r}
descr = describe(PNC$age[PNC$sex == 'MALE'])[,c(2:5, 8:9),]
row.names(descr) = "age"
descr %>% pander()
```

Females:
```{r}
descr = describe(PNC$age[PNC$sex == 'FEMALE'])[,c(2:5, 8:9),]
row.names(descr) = "age"
descr %>% pander()
```

## ICV: Models by Sex

```{r, results='asis'}
lm_M = lm(scale(ICV) ~ scale(age)  , data = filter(PNC, sex == "MALE"))
lm_F = lm(scale(ICV) ~ scale(age)  , data = filter(PNC, sex == "FEMALE"))

# Show table (pretty)
#suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Intracranial Volume ",
          covariate.labels = c("Age"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "0pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

## White Matter: Models by Sex


```{r, results='asis'}
lm_M = lm(scale(WM) ~ scale(age)  , data = filter(PNC, sex == "MALE"))
lm_F = lm(scale(WM) ~ scale(age)  , data = filter(PNC, sex == "FEMALE"))

# Show table (pretty)
# suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "White Matter Volume ",
          covariate.labels = c("Age"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "0pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

## Gray Matter: Models by Sex

```{r, results='asis'}
lm_M = lm(scale(GM) ~ scale(age)  , data = filter(PNC, sex == "MALE"))
lm_F = lm(scale(GM) ~ scale(age)  , data = filter(PNC, sex == "FEMALE"))

# Show table (pretty)
# suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Gray Matter Volume ",
          covariate.labels = c("Age"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "0pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

## CSF (in ventricles): Models by Sex

```{r, results='asis'}
lm_M = lm(scale(CSF) ~ scale(age)  , data = filter(PNC, sex == "MALE"))
lm_F = lm(scale(CSF) ~ scale(age)  , data = filter(PNC, sex == "FEMALE"))

# Show table (pretty)
# suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Cerebrospinal Fluid Volume ",
          covariate.labels = c("Age"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "0pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

## Comparison of GAMs vs Linear models across all 145 ROIs

- Model 1: ROI = sex + age + age*sex   
- Model 2: ROI = sex + s(age, by = sex), where s() denotes the smooth function(s) fitted separately within each sex.


## ROIs w/ non-linear age trends (FDR-adjusted): 

```{r}
## ANOVA of all regions
rois = PNC[, roi_small]
Ps = matrix(nrow=2, ncol=ncol(rois))
colnames(Ps) = colnames(rois)

# Linear vs GAM
l = rois %>% lapply(function(x) anova(lm(x ~ sex + age + age*sex, data = PNC),
                                      gam(x ~ s(age, by = sex, bs = 'tp', k=3, fx=TRUE) + sex,
                                          data = PNC, family = gaussian, method = 'REML')))

for(j in colnames(rois)) Ps[1,j] = l[[j]]$`Pr(>F)`[2]

p = c(sum(Ps[1,] < 0.05),
      sum(p.adjust(Ps[1,], method = 'bonferroni') < 0.05),
      sum(p.adjust(Ps[1,], method = 'fdr') < 0.05))
names(p) = c("GAM != Linear (No adj.)", "GAM != Linear (Bonferroni)", "GAM != Linear (FDR)")

sig_rois = names(which(p.adjust(Ps[1,], method = 'fdr') < 0.05))
dict[dict$ROI_INDEX %in% readr::parse_number(sig_rois),1:4] %>% pander()
```

## Pairwise Correlations

```{r, warning=FALSE, message=FALSE}
sig_rois_paired = matrix(data = NA, nrow = nrow(PNC), ncol = length(sig_rois) %/% 2)

for(i in 2:length(sig_rois)){
  if(i %% 2 != 0) next # skip odd numbers
  sig_rois_paired[,i/2] =  rowSums(PNC[,sig_rois[c(i, i+1)]])
}

sig_rois_df = data.frame(PNC[,c("sex", "X35")], sig_rois_paired)
colnames(sig_rois_df) = c("sex", "Brain Stem", "Cerebellum WM", "Ventral DC", "Fornix", "Ant Limb Int Capsule")

df.m = melt(sig_rois_df)

upper.panel<-function(x, y){
  points(x,y, pch=19, col=c("red", "blue")[sig_rois_df$sex])
  r <- round(cor(x, y), digits=2)
  txt <- paste0("R = ", r)
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  text(0.5, 0.9, txt)
}
pairs(sig_rois_df[,-1], lower.panel = NULL, 
      upper.panel = upper.panel, main = "Pairwise correlations of significant ROIs (GAM)") %>% suppressWarnings()
```

## Brain Stem: GAM Plot

```{r}
sig_rois_df$age = PNC$age # add age
colnames(sig_rois_df) = gsub(" ", "\\.", colnames(sig_rois_df)) # replace spaces for dots in column names

# Males
gam_M = mgcv::gam(Brain.Stem ~ s(age,
                                  k=3,
                                  bs = 'tp',
                                  fx=TRUE), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')

# Females
gam_F = mgcv::gam(Brain.Stem ~ s(age,
                                k=3,
                                bs = 'tp',
                                fx=TRUE), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, main = "Females", ylab = "Brain Stem", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1,  main = 'Males', cex = 0.75, ylab = "Brain Stem", shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

## Brain Stem: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Brain.Stem) ~ scale(age) + scale((age - mean(age))^2), data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Brain.Stem) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
#suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Brain Stem Volume",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "0pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

## Cerebellum WM: GAM plot

```{r}
# Males
gam_M = mgcv::gam(Cerebellum.WM ~ s(age,
                                  k=3,
                                  bs = 'tp', 
                                  fx=TRUE), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(Cerebellum.WM ~ s(age,
                                k=3,
                                bs = 'tp',
                                fx=TRUE), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "Cerebellum WM", main = "Females", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1, ylab = "Cerebellum WM", main = 'Males', cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

## Cerebellum WM: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Cerebellum.WM) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Cerebellum.WM) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Cerebellum WM ",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "0pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

R-squared of GAMs:  
- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 


## Ventral Diencephalon: GAM Plot

```{r}
# Males
gam_M = mgcv::gam(Ventral.DC ~ s(age,
                                  k=3,
                                  bs = 'tp', 
                                  fx=TRUE), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(Ventral.DC ~ s(age,
                                k=3,
                                bs = 'tp',
                                fx=TRUE), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "Ventral DC", main = "Females", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1, ylab = "Ventral DC", main = 'Males', cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

## Ventral DC: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Ventral.DC) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Ventral.DC) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Ventral DC",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "0pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

## Fornix: GAM Plot

```{r}
# Males
gam_M = mgcv::gam(Fornix ~ s(age,
                                  k=3,
                                  bs = 'tp', 
                                  fx=TRUE), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(Fornix ~ s(age,
                                k=3,
                                bs = 'tp',
                                fx=TRUE), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "Fornix", main = "Females", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1, ylab = "Fornix", main = 'Males', cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

## Fornix: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Fornix) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Fornix) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Fornix",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "0pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

## Anterior Limb of Internal Capsule (ALIC): GAM Plot

```{r}
# Males
gam_M = mgcv::gam(Ant.Limb.Int.Capsule ~ s(age,
                                  k=3,
                                  bs = 'tp', 
                                  fx=TRUE), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(Ant.Limb.Int.Capsule ~ s(age,
                                k=3,
                                bs = 'tp',
                                fx=TRUE), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "Ant. Limb Internal Capsule", main = "Females", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1, ylab = "Ant. Limb Internal Capsule", main = 'Males', cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

## ALIC: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Ant.Limb.Int.Capsule) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Ant.Limb.Int.Capsule) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Ant. Limb Internal Capsule",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "0pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

# Harmonization

## Factors guiding harmonization approach

- Differential Age confound in Males vs. Females
- ICV, ROI volume differences in Males vs. Females
- Differential relationships (e.g., ROI-age, ROI-ICV) in Males vs. Females
- Non-linear age trends in ROIs

## Harmonization Approach

Adjusted data are shown for the following approach:  

- Join HC and MS data into one dataset
- Split this dataset into 2 according to sex.
- Harmonize (Step 1): Run ComBat-GAM with ICV as target.
- Harmonize (Step 1): Run ComBat-GAM with 145 as target with harmonized ICV as added covariate.

Models used in harmonization

- GAM: s(age) + MS + MS*x*age

```{r}
check_roi_site_fx = function(df, roi_names, mod_sex = FALSE, mod_MS = FALSE, mod_interaction = TRUE, print_Ps = FALSE){
   base_rh_formula = "~ ICV + age + age2"
   
   if(mod_sex){
      if(mod_interaction){
         sex_terms = " + sex + sex*age + sex*age2"
      }else{
         sex_terms = " + sex"
      }
   }else{
      sex_terms = ""
   }
   
   if(mod_MS){
      if(mod_interaction){
         ms_terms = " + MS + MS*age + MS*age2"
      }else{
         ms_terms = " + MS"
      }
         
   }else{
      ms_terms = ""
   }
   
   # Decide formula 
   rh_formula = paste0(base_rh_formula, sex_terms, ms_terms)
   print(rh_formula)
   extract_p = function(x) anova(lm(as.formula(paste0(x, rh_formula)), data = df), 
                                 lm(as.formula(paste0(x, rh_formula, "+ site")), data = df)
                                       )$`Pr(>F)`[2]
      
   
   Ps = roi_names %>% sapply(extract_p)
   p_count = c(sum(p.adjust(Ps, method = 'fdr') < 0.05),
               sum(p.adjust(Ps, method = 'bonferroni') < 0.05),
               sum(Ps < 0.05)
               )
   if(print_Ps) print(Ps)
   names(p_count) = c("FDR", "Bonferroni", "Uncorrected P")
   invisible(p_count)
}
```

## Testing for site effects

ANOVAs were run on each of the 145 ROIs comparing two versions of covariate model: with and without the inclusion of a site/scanner variable.


## Site effects: MS + HC 

Number of ROIs showing site effects:

Raw data:

```{r}
# Full model
rbind(HC_raw, MS_raw) %>% check_roi_site_fx(roi_small, mod_sex = TRUE, mod_MS = TRUE) %>% pander()
```

Harmonized data: 

```{r}
# Full model
rbind(HC_GAM, MS_GAM) %>% check_roi_site_fx(roi_small, mod_sex = TRUE, mod_MS = TRUE) %>% pander()
```

## Site effects: MS + HC [Females]

Raw:

```{r}
rbind(HC_raw, MS_raw) %>%  filter(sex == "FEMALE") %>% check_roi_site_fx(roi_small, mod_MS = TRUE) %>% pander()
```

Harmonized:

```{r}
rbind(HC_GAM, MS_GAM) %>%  filter(sex == "FEMALE") %>% check_roi_site_fx(roi_small, mod_MS = TRUE) %>% pander()
```

## Site effects: MS + HC [Males]

Raw:

```{r}
rbind(HC_raw, MS_raw) %>% filter(sex == "MALE") %>%  check_roi_site_fx(roi_small, mod_MS = TRUE) %>% pander()
```

Harmonized:
```{r}
rbind(HC_GAM, MS_GAM) %>% filter(sex == "MALE") %>%  check_roi_site_fx(roi_small, mod_MS = TRUE) %>% pander()
```

## Site effects: MS 

Raw:

```{r}
# Full model
MS_raw %>% check_roi_site_fx(roi_small, mod_sex = TRUE) %>% pander() 
```

Harmonized:
```{r}
MS_GAM %>% check_roi_site_fx(roi_small, mod_sex = TRUE) %>% pander() 
```


## Site effects: MS [Females]

Raw:
```{r}
MS_raw %>% filter(sex == "FEMALE") %>%  check_roi_site_fx(roi_small, mod_sex = FALSE) %>% pander()
```

Harmonized:

```{r}
MS_GAM %>% filter(sex == "FEMALE") %>%  check_roi_site_fx(roi_small, mod_sex = FALSE) %>% pander()
```


## Site effects: MS [Males]

Raw:   
```{r}
MS_GAM %>% filter(sex == "MALE") %>%  check_roi_site_fx(roi_small, mod_sex = FALSE) %>% pander()
```

Harmonized:  
```{r}
MS_GAM %>% filter(sex == "MALE") %>%  check_roi_site_fx(roi_small, mod_sex = FALSE) %>% pander()
```

## Site effects: HC

Raw: 

```{r}
# Full model
HC_raw %>% check_roi_site_fx(roi_small, mod_sex = TRUE) %>% pander() 
```

Harmonized: 

```{r}
# Full model
HC_GAM %>% check_roi_site_fx(roi_small, mod_sex = TRUE) %>% pander() 
```

## Site effect: HC [Females]

Raw

```{r}
HC_raw %>% filter(sex == "FEMALE") %>%  check_roi_site_fx(roi_small, mod_sex = FALSE) %>% pander()
```

Harmonized:

```{r}
HC_GAM %>% filter(sex == "FEMALE") %>%  check_roi_site_fx(roi_small, mod_sex = FALSE) %>% pander()
```

## Site effects: HC [Males]

Raw:
```{r}
HC_raw %>% filter(sex == "MALE") %>%  check_roi_site_fx(roi_small, mod_sex = FALSE) %>% pander()
```

Harmonized:
```{r}
HC_GAM %>% filter(sex == "MALE") %>%  check_roi_site_fx(roi_small, mod_sex = FALSE) %>% pander()
```

## Harmonization Conclusions

Insert conclusions






