---
title: 'EDA: GAMs'
author: "Virgilio Gonzenbach"
date: "1/13/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      cache = TRUE, 
                      cache.path = 'cache/GAMs/', 
                      fig.path = "figures/GAMs/")
knitr::opts_knit$set(root.dir = '../')
```

```{r, include=FALSE}
library(mgcv)
library(readxl)
library(tidyverse) # dplyr and ggplot
library(psych)
library(stargazer)
library(pander)
library(ggpubr)
library(reshape2)

df = read.csv("data/deriv/HC_data.csv", stringsAsFactors = TRUE)
dict = read_xlsx("data/MUSE_ROI_Dict.xlsx", 1)

# Improve readibility for raw ICV
colnames(df)[6] = "raw.ICV"

# Shorten Dictionary
dict = dict[dict$ROI_INDEX %in% 2:300, ]

# Tissue filters
VN_filter = dict$TISSUE_SEG == "VN"
GM_filter = dict$TISSUE_SEG == "GM"
WM_filter = dict$TISSUE_SEG == "WM"

# Select PNC dataset
pnc = df %>% filter(site == "PNC") 
pnc$age2 =  age - mean(age)^2

# Aggregate ROIs
rois = pnc[, 7:ncol(pnc)]
pnc$WM = rowSums(rois[, WM_filter])
pnc$GM = rowSums(rois[, GM_filter])
pnc$CSF = rowSums(rois[, VN_filter])
pnc$BrStem = pnc$X35

# Separate datasets by sex
pnc_M = filter(pnc, sex == "MALE")
pnc_F = filter(pnc, sex == "FEMALE")

# Parameters for GAMs
K = 3
FX = TRUE
```


## Objective

To inform the potential usage of a nonlinear method of harmonization (ComBat-GAM) by:

1. Visualizing nonlinear effects of age on various volumetric regions of interest (ROIs),
2. Modeling these nonlinear effects using Generative Additive Models (GAMs), and 
3. Making comparisons between GAMs and their linear counter parts.

All modeling and visualization carried out in PNC data.

Note: All GAMs constrained to K = `r K` and FX = `r FX`, where K denotes the number of basis functions and FX indicates the use of fixed degrees of freedom. 

\pagebreak 

## Age Descriptives

```{r}
p = df %>% ggplot(aes(x=sex, y=age, color = sex)) + geom_boxplot() + ggtitle("Distribution of Age by Sex in PNC data") + ylab("Age") + xlab("Sex") + theme(plot.title = element_text(hjust = 0.5))
change_palette(p, palette = "npg")
```

Full dataset:   
```{r}
descr = describe(pnc$age)[,c(2:5, 8:9),]
row.names(descr) = "age"
descr %>% pander()
```

Males:
```{r}
library(psych)
descr = describe(pnc$age[pnc$sex == 'MALE'])[,c(2:5, 8:9),]
row.names(descr) = "age"
descr %>% pander()
```

Females:
```{r}
library(psych)
descr = describe(pnc$age[pnc$sex == 'FEMALE'])[,c(2:5, 8:9),]
row.names(descr) = "age"
descr %>% pander()
```

\pagebreak

## Comparison of GAMs vs Linear models across all 145 ROIs

-Model 1: ROI = sex + age + age*sex  
-Model 2: ROI = sex + s(age, by = sex), where s() denotes the smooth function(s) fitted separately within each sex.

Number of regions for which models (GAM and linear model) significantly differed:  

```{r}
## ANOVA of all regions
Ps = matrix(nrow=2, ncol=ncol(rois))
colnames(Ps) = colnames(rois)

# Linear vs GAM
l = rois %>% lapply(function(x) anova(lm(x ~ sex + age + age*sex, data = pnc),
                                      gam(x ~ s(age, by = sex, bs = 'tp', k=3, fx=FX) + sex, 
                                          data = pnc, family = gaussian, method = 'REML')))

for(j in colnames(rois)) Ps[1,j] = l[[j]]$`Pr(>F)`[2]

p = c(sum(Ps[1,] < 0.05),
      sum(p.adjust(Ps[1,], method = 'bonferroni') < 0.05),
      sum(p.adjust(Ps[1,], method = 'fdr') < 0.05))
names(p) = c("GAM != Linear (No adj.)", "GAM != Linear (Bonferroni)", "GAM != Linear (FDR)")
p
```

ROIs where GAM differed from linear model (adjusted for False Discovery Rate): 

```{r}
sig_rois = names(which(p.adjust(Ps[1,], method = 'fdr') < 0.05))
dict[dict$ROI_INDEX %in% parse_number(sig_rois),1:4] %>% pander()
```

## Comparison Linear models with and without quadratic term (within each sex)

-Model 1: $ROI = age + age^2$
-Model 2: $ROI = sex + s(age, by = sex)$, where s() denotes the smooth function(s) fitted separately within each sex.

### Males

Number of regions for which models (GAM and linear model) significantly differed:  

```{r}
## ANOVA of all regions
Ps = matrix(nrow=2, ncol=ncol(rois))
colnames(Ps) = colnames(rois)

# Linear vs GAM
l = rois %>% lapply(function(x){} anova(lm(x ~ age, data = filter(pnc, sex == "MALE")),
                                      lm(x ~ age + (age - mean(age))^2, data = filter(pnc, sex == "MALE"))
                                      )
)

for(j in colnames(rois)) Ps[1,j] = l[[j]]$`Pr(>F)`[2]

p = c(sum(Ps[1,] < 0.05),
      sum(p.adjust(Ps[1,], method = 'bonferroni') < 0.05),
      sum(p.adjust(Ps[1,], method = 'fdr') < 0.05))
names(p) = c("GAM != Linear (No adj.)", "GAM != Linear (Bonferroni)", "GAM != Linear (FDR)")
p
```

ROIs where quadratic term improved the model (adjusted for False Discovery Rate): 

```{r}
sig_rois = names(which(p.adjust(Ps[1,], method = 'fdr') < 0.05))
dict[dict$ROI_INDEX %in% parse_number(sig_rois),1:4] %>% pander()
```

## Regions of Interest (Significant)

```{r, include=FALSE}
## Box plots

sig_rois_paired = matrix(data = NA, nrow = nrow(pnc), ncol = length(sig_rois) %/% 2)

for(i in 2:length(sig_rois)){
  if(i %% 2 != 0) next # skip odd numbers
  sig_rois_paired[,i/2] =  rowSums(pnc[,sig_rois[c(i, i+1)]])
}

sig_rois_df = data.frame(pnc[,c("sex", "X35")], sig_rois_paired)
colnames(sig_rois_df) = c("sex", "Brain Stem", "Cerebellum WM", "Ventral DC", "Fornix", "Ant Limb Int Capsule")

df.m = melt(sig_rois_df)
```


```{r}
df.m %>% ggplot(aes(x=variable, y=value, col=sex)) + geom_boxplot() + ggtitle("Significant GAMs") + ylab("Volume") + xlab("Region of Interest")

## change palette
change_palette(p, "npg")
```


```{r}
upper.panel<-function(x, y){
  points(x,y, pch=19, col=c("red", "blue")[sig_rois_df$sex])
  r <- round(cor(x, y), digits=2)
  txt <- paste0("R = ", r)
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  text(0.5, 0.9, txt)
}
pairs(sig_rois_df[,-1], lower.panel = NULL, 
      upper.panel = upper.panel, main = "Pairwise correlations of significant ROIs (GAM)")
```

\pagebreak

## Brain Stem
### Brain Stem: Plot GAM

```{r}
sig_rois_df$age = pnc$age # add age
colnames(sig_rois_df) = gsub(" ", "\\.", colnames(sig_rois_df)) # replace spaces for dots in column names

# Males
gam_M = mgcv::gam(Brain.Stem ~ s(age,
                                  k=K,
                                  bs = 'tp',
                                  fx=FX), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')

# Females
gam_F = mgcv::gam(Brain.Stem ~ s(age,
                                k=K,
                                bs = 'tp',
                                fx=FX), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, main = "Females", ylab = "Brain Stem", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1,  main = 'Males', cex = 0.75, ylab = "Brain Stem", shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

### Brain Stem: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Brain.Stem) ~ scale(age) + scale((age - mean(age))^2), data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Brain.Stem) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
#suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Brain Stem Volume",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "25pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

\pagebreak

### Brain Stem: GAM summary

```{r}
gam = mgcv::gam(Brain.Stem ~ s(age,  by = sex, bs = 'tp', k=K, fx=FX) + sex, 
                    data = sig_rois_df,
                    family = gaussian,
                    method = 'REML')
summary(gam)
```

Compare with R-squared of the equivalent linear model (ICV ~ sex + age + age *x* sex): `r round(summary(lm(Brain.Stem ~ scale(age) + sex*scale(age), data = sig_rois_df))$r.squared, 3)`

Compare GAM and linear model directly:

```{r}
anova(lm(Brain.Stem~ scale(age) + sex*scale(age) + sex, data = sig_rois_df), gam)
```

\pagebreak

## Cerebellum White Matter
### Cerebellum WM: Plot GAM

```{r}
# Males
gam_M = mgcv::gam(Cerebellum.WM ~ s(age,
                                  k=K,
                                  bs = 'tp', 
                                  fx=FX), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(Cerebellum.WM ~ s(age,
                                k=K,
                                bs = 'tp',
                                fx=FX), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "Cerebellum WM", main = "Females", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1, ylab = "Cerebellum WM", main = 'Males', cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

### Cerebellum WM: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Cerebellum.WM) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Cerebellum.WM) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Cerebellum WM ",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "25pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

R-squared of GAMs:  
- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

\pagebreak

### Cerebellum WM: GAM summary

```{r}
gam = mgcv::gam(Cerebellum.WM ~ s(age,  by = sex, bs = 'tp', k=K, fx=FX) + sex, 
                    data = sig_rois_df,
                    family = gaussian,
                    method = 'REML')
summary(gam)
```

Compare with R-squared of the equivalent linear model (Cerebellum.WM ~ sex + age + age *x* sex): `r round(summary(lm(Cerebellum.WM ~ scale(age) + sex*scale(age), data = sig_rois_df))$r.squared, 3)`

Compare GAM and linear model directly:

```{r}
anova(lm(Cerebellum.WM ~ scale(age) + sex*scale(age) + sex, data = sig_rois_df), gam)
```



\pagebreak

## Ventral Diencephalon
### Ventral DC: Plot GAM

```{r}
# Males
gam_M = mgcv::gam(Ventral.DC ~ s(age,
                                  k=K,
                                  bs = 'tp', 
                                  fx=FX), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(Ventral.DC ~ s(age,
                                k=K,
                                bs = 'tp',
                                fx=FX), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "Ventral DC", main = "Females", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1, ylab = "Ventral DC", main = 'Males', cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

### Ventral DC: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Ventral.DC) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Ventral.DC) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Ventral DC",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "25pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

\pagebreak

### Ventral DC: GAM summary

```{r}
gam = mgcv::gam(Ventral.DC ~ s(age,  by = sex, bs = 'tp', k=K, fx=FX) + sex, 
                    data = sig_rois_df,
                    family = gaussian,
                    method = 'REML')
summary(gam)
```

Compare with R-squared of the equivalent linear model (Ventral.DC ~ sex + age + age *x* sex): `r round(summary(lm(Ventral.DC ~ scale(age) + sex*scale(age), data = sig_rois_df))$r.squared, 3)`

Compare GAM and linear model directly:

```{r}
anova(lm(Ventral.DC ~ scale(age) + sex*scale(age) + sex, data = sig_rois_df), gam)
```

\pagebreak

## Fornix
### Fornix: Plot GAM

```{r}
# Males
gam_M = mgcv::gam(Fornix ~ s(age,
                                  k=K,
                                  bs = 'tp', 
                                  fx=FX), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(Fornix ~ s(age,
                                k=K,
                                bs = 'tp',
                                fx=FX), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "Fornix", main = "Females", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1, ylab = "Fornix", main = 'Males', cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

### Fornix: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Fornix) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Fornix) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Fornix",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "25pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

\pagebreak

### Fornix: GAM summary

```{r}
gam = mgcv::gam(Fornix ~ s(age,  by = sex, bs = 'tp', k=K, fx=FX) + sex, 
                    data = sig_rois_df,
                    family = gaussian,
                    method = 'REML')
summary(gam)
```

Compare with R-squared of the equivalent linear model (Fornix ~ sex + age + age *x* sex): `r round(summary(lm(Fornix ~ scale(age) + sex*scale(age), data = sig_rois_df))$r.squared, 3)`

Compare GAM and linear model directly:

```{r}
anova(lm(Fornix ~ scale(age) + sex*scale(age) + sex, data = sig_rois_df), gam)
```

\pagebreak

## Anterior Limb of Internal Capsule (ALIC)
### ALIC: Plot GAM

```{r}
# Males
gam_M = mgcv::gam(Ant.Limb.Int.Capsule ~ s(age,
                                  k=K,
                                  bs = 'tp', 
                                  fx=FX), 
                    data = filter(sig_rois_df, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(Ant.Limb.Int.Capsule ~ s(age,
                                k=K,
                                bs = 'tp',
                                fx=FX), 
                    data = filter(sig_rois_df, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "Ant. Limb Internal Capsule", main = "Females", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1, ylab = "Ant. Limb Internal Capsule", main = 'Males', cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1])
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

### ALIC: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(Ant.Limb.Int.Capsule) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "MALE"))
lm_F = lm(scale(Ant.Limb.Int.Capsule) ~ scale(age) + scale((age - mean(age))^2) , data = filter(sig_rois_df, sex == "FEMALE"))

# Show table (pretty)
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Ant. Limb Internal Capsule",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "25pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

\pagebreak
### ALIC: GAM summary

```{r}
gam = mgcv::gam(Ant.Limb.Int.Capsule ~ s(age,  by = sex, bs = 'tp', k=K, fx=FX) + sex, 
                    data = sig_rois_df,
                    family = gaussian,
                    method = 'REML')
summary(gam)
```

Compare with R-squared of the equivalent linear model (ALIC ~ sex + age + age *x* sex): `r round(summary(lm(Ant.Limb.Int.Capsule ~ scale(age) + sex*scale(age), data = sig_rois_df))$r.squared, 3)`

Compare GAM and linear model directly:

```{r}
anova(lm(Ant.Limb.Int.Capsule ~ scale(age) + sex*scale(age) + sex, data = sig_rois_df), gam)
```


\pagebreak

## Intracranial Volume
### ICV: Plot GAM

```{r}
# Males
gam_M = mgcv::gam(raw.ICV ~ s(age,
                                  k=K,
                                  bs = 'tp', 
                                  fx=FX), 
                    data = filter(pnc, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(raw.ICV ~ s(age,
                                k=K,
                                bs = 'tp',
                                fx=FX), 
                    data = filter(pnc, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, main = "Females", ylab = "raw ICV", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1])
plot(gam_M, residuals = TRUE, pch = 1, cex = 0.75, ylab = "raw ICV", shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1], main = 'Males')
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

### ICV: Linear Models by Sex

```{r, results='asis'}
lm_M = lm(scale(raw.ICV) ~ scale(age) + scale((age - mean(age))^2) , data = filter(pnc, sex == "MALE"))
lm_F = lm(scale(raw.ICV) ~ scale(age) + scale((age - mean(age))^2) , data = filter(pnc, sex == "FEMALE"))

# Show table (pretty)
#suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Intracranial Volume ",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "25pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

\pagebreak

### ICV: GAM summary

```{r}
gam = mgcv::gam(raw.ICV ~ s(age,  by = sex, bs = 'tp', k=K, fx=FX) + sex, 
                    data = pnc,
                    family = gaussian,
                    method = 'REML')
summary(gam)
```

Model indicates ICV-to-Age relationship in males is approx. linear (EDF $\approx$ 1) and significant. In females, however, the relationship is closer to a quadratic term (EDF  2) and non-significant.

Compare with R-squared of the equivalent linear model (ICV ~ sex + age + age *x* sex): `r round(summary(lm(raw.ICV ~ scale(age) + sex*scale(age), data = pnc))$r.squared, 3)`

Compare GAM and linear model directly:

```{r}
anova(lm(raw.ICV ~ scale(age) + sex*scale(age) + sex, data = pnc), gam)
```

\pagebreak

## White Matter
### WM: Plot GAMs

```{r}
# Males
gam_M = mgcv::gam(WM ~ s(age,
                            k=K, fx=FX,
                            bs = 'tp'), 
                    data = filter(pnc, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(WM ~ s(age,
                            k=K, fx=FX,
                            bs = 'tp'), 
                    data = filter(pnc, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "raw WM", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1], main = "Females")
plot(gam_M, residuals = TRUE, pch = 1, ylab = "raw WM", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1], main = 'Males')
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

### WM: Linear models by sex

```{r, results='asis'}
lm_M = lm(scale(WM) ~ scale(age) + scale((age - mean(age))^2) , data = filter(pnc, sex == "MALE"))
lm_F = lm(scale(WM) ~ scale(age) + scale((age - mean(age))^2) , data = filter(pnc, sex == "FEMALE"))

# Show table (pretty)
# suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "White Matter Volume ",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "25pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

\pagebreak

### WM: GAM summary

```{r}
gam = mgcv::gam(WM ~ s(age,  by = sex, bs = 'tp', k=K, fx=FX) + sex, 
                    data = pnc,
                    family = gaussian,
                    method = 'REML')
summary(gam)
```

Model indicates WM-to-Age relationship is nonlinear and statistically significant in both males and females. WM increases with age; but in females, growth stops at 20 years of age.

Compare with R-squared of the equivalent linear model (WM ~ sex + age + age *x* sex): `r round(summary(lm(WM ~ scale(age) + sex*scale(age), data = pnc))$adj.r.squared, 3)`

Compare GAM and linear model directly:

```{r}
anova(lm(WM ~ scale(age) + sex*scale(age) + sex, data = pnc), gam)
```

\pagebreak

## Gray Matter
### GM: Plot GAMs

```{r}
# Males
gam_M = mgcv::gam(GM ~ s(age,
                                k=K, fx=FX,
                                bs = 'tp'), 
                    data = filter(pnc, sex == "MALE"),
                    family = gaussian,
                    method = 'REML')
# Females
gam_F = mgcv::gam(GM ~ s(age,
                                k=K, fx=FX,
                                bs = 'tp'), 
                    data = filter(pnc, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow= c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "raw GM", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1], main = "Females")
plot(gam_M, residuals = TRUE, pch = 1, ylab = "raw GM", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1], main = 'Males')
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

### GM: Linear models by sex

```{r, results='asis'}
lm_M = lm(scale(GM) ~ scale(age) + scale((age - mean(age))^2) , data = filter(pnc, sex == "MALE"))
lm_F = lm(scale(GM) ~ scale(age) + scale((age - mean(age))^2) , data = filter(pnc, sex == "FEMALE"))

# Show table (pretty)
# suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Gray Matter Volume ",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "25pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

### GM: GAM summary

```{r}
gam = mgcv::gam(GM ~ s(age,  by = sex, bs = 'tp', k=K, fx=FX) + sex, 
                    data = pnc,
                    family = gaussian,
                    method = 'REML')
summary(gam)
```

Model indicates GM-to-age relationship is linear and statistically significant for both males and females. GM decreases with age.

Compare with adj. R-squared of the equivalent linear model (WM ~ sex + age + age *x* sex): `r round(summary(lm(GM ~ scale(age) + sex*scale(age), data = pnc))$adj.r.squared, 3)`

Compare GAM and linear model directly:

```{r}
anova(lm(GM ~ scale(age) + sex*scale(age) + sex, data = pnc), gam)
```

\pagebreak

## Cerebrospinal Fluid (in ventricles only)
### CSF: Plot GAMs

```{r}
# Males
gam_M = mgcv::gam(CSF ~ s(age,
                             bs = 'tp', k=K, fx=FX), 
                    data = pnc[pnc$sex == "MALE",],
                    family = gaussian,
                    method = 'REML')
```

```{r}
# Females
gam_F = mgcv::gam(CSF ~ s(age,
                             bs = 'tp', k=K, fx=FX), 
                    data = filter(pnc, sex == "FEMALE"),
                    family = gaussian,
                    method = 'REML')

par(mfrow = c(1, 2))
plot(gam_F, residuals = TRUE, pch = 1, ylab = "raw CSF", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_F)[1], main = "Females")
plot(gam_M, residuals = TRUE, pch = 1, ylab = "raw CSF", cex = 0.75, shade = TRUE, seWithMean = TRUE, shift = coef(gam_M)[1], main = 'Males')
```

```{r}
summ_M = summary(gam_M)
summ_F = summary(gam_F)
```

Compare with R-squared of GAM:  
- Males: $R^2 =$ `r round(summ_M$r.sq,3)`  
- Females: $R^2 =$ `r round(summ_F$r.sq,3)` 

### CSF: Linear models by sex

```{r, results='asis'}
lm_M = lm(scale(CSF) ~ scale(age) + scale((age - mean(age))^2) , data = filter(pnc, sex == "MALE"))
lm_F = lm(scale(CSF) ~ scale(age) + scale((age - mean(age))^2) , data = filter(pnc, sex == "FEMALE"))

# Show table (pretty)
# suppressMessages(tab_model(lm_M, lm_F,p.style = "numeric"))
stargazer(lm_M, lm_F, 
          column.labels = c("Males", "Females"),
          dep.var.labels = "Cerebrospinal Fluid Volume ",
          covariate.labels = c("Age", "Age (Squared)"),
          keep.stat = c("rsq"),
          notes = "Stargazer", 
          header=FALSE,
          column.sep.width = "25pt",
          single.row = TRUE,
          ci = TRUE,
          star.cutoffs = c(.05, .01, .001)
          )
```

\pagebreak

### CSF: GAM summary

```{r}
gam = mgcv::gam(CSF ~ s(age,  by = sex, bs = 'tp', k=K, fx=FX) + sex, 
                    data = pnc,
                    family = gaussian,
                    method = 'REML')
summary(gam)
```

Model indicates CSF-to-age relationship is linear and statistically significant in males and females. CSF increases with age.

Compare with adj. R-squared of the equivalent linear model (CSF ~ sex + age + age *x* sex): `r round(summary(lm(CSF ~ scale(age) + sex*scale(age), data = pnc))$adj.r.squared, 4)`

Compare GAM and linear model directly:

```{r}
anova(lm(CSF ~ scale(age) + sex*scale(age) + sex, data = pnc), gam)
```
